---
title: "deep_learning_example-1"
author: "Ashish Dutt"
date: "February 7, 2018"
output: pdf_document
---
The following example is taken from the website (https://keras.rstudio.com/)
```{r}
# clean the workspace
rm(list = ls())
#install.packages("keras", dependencies = TRUE)
library(keras)
# Read in MNIST data
mnist <- dataset_mnist()

x_train <- mnist$train$x
y_train <- mnist$train$y
x_test <- mnist$test$x
y_test <- mnist$test$y
```
The x data is a 3-d array (images,width,height) of grayscale values . To prepare the data for training we convert the 3-d arrays into matrices by reshaping width and height into a single dimension (28x28 images are flattened into length 784 vectors). Then, we convert the grayscale values from integers ranging between 0 to 255 into floating point values ranging between 0 and 1:
```{r}
# reshape
x_train <- array_reshape(x_train, c(nrow(x_train), 784))
x_test <- array_reshape(x_test, c(nrow(x_test), 784))
# rescale
x_train <- x_train / 255
x_test <- x_test / 255
```
Note that we use the `array_reshape()` function rather than the dim<-() function to reshape the array. This is so that the data is re-interpreted using row-major semantics (as opposed to R’s default column-major semantics), which is in turn compatible with the way that the numerical libraries called by Keras interpret array dimensions.

The y data is an integer vector with values ranging from 0 to 9. To prepare this data for training we one-hot encode the vectors into binary class matrices using the Keras `to_categorical()` function

```{r}
y_train <- to_categorical(y_train, 10)
y_test <- to_categorical(y_test, 10)
```
#### Defining the Model
The core data structure of Keras is a model, a way to organize layers. The simplest type of model is the Sequential model, a linear stack of layers.

We begin by creating a sequential model and then adding layers using the pipe (%>%) operator:
```{r}
model <- keras_model_sequential() 
model %>% 
  layer_dense(units = 256, activation = 'relu', input_shape = c(784)) %>% 
  layer_dropout(rate = 0.4) %>% 
  layer_dense(units = 128, activation = 'relu') %>%
  layer_dropout(rate = 0.3) %>%
  layer_dense(units = 10, activation = 'softmax')
```
The input_shape argument to the first layer specifies the shape of the input data (a length 784 numeric vector representing a grayscale image). The final layer outputs a length 10 numeric vector (probabilities for each digit) using a softmax activation function.

Use the `summary()` function to print the details of the model:
```{r}
# Print a summary of the model
summary(model)
# Get model configuration
get_config(model)

# Get layer configuration
get_layer(model, index = 1)

# List the model's layers
model$layers

# List the input tensors
model$inputs

# List the output tensors
model$outputs
```
### Compile the Model

```{r}
model %>% compile(
  loss = 'categorical_crossentropy',
  optimizer = optimizer_rmsprop(),
  metrics = c('accuracy')
)

```
### Training and Evaluation
Use the fit() function to train the model for 30 epochs using batches of 128 images:

```{r}
history <- model %>% fit(
  x_train, y_train, 
  epochs = 30, batch_size = 128, 
  validation_split = 0.2
)
```
### Plot the Model
The `history` object returned by `fit()` includes loss and accuracy metrics which we can plot:
```{r}
plot(history)
```
### Evaluating model performance
Evaluate the model’s performance on the test data:
```{r}
model %>% evaluate(x_test, y_test)
```
Generate predictions on new data:
```{r}
model %>% predict_classes(x_test)
```