---
title: "How to split a dataset into train and test"
author: "Ashish Dutt"
date: "February 6, 2018"
output: pdf_document
---

## Data Splitting

There are several approaches to splitiing a dataset into train and test for a predictive modelling task. I present a few different approaches below;

### Approach# 1: Manual approach: split into training and test sets having a 60/40 split 
```{r}
set.seed(2018)
train_indices <- 1:round(0.6 * n)
train <- shuffled_df[train_indices, ]
test_indices <- (round(0.6 * n) + 1):n
test <- shuffled_df[test_indices, ]
dim(train)
dim(test)
```

### Approach# 2: using the sample() method
```{r}
data(mtcars)
dim(mtcars)
```
#### 75% of the sample size
```{r}
smp_size <- floor(0.75 * nrow(mtcars))
```
#### set the seed to make your partition reproductible
```{r}
set.seed(2018)
train_ind <- sample(seq_len(nrow(mtcars)), size = smp_size)
train <- mtcars[train_ind, ]
test <- mtcars[-train_ind, ]
dim(train)
dim(test)
```

### Approach# 3: using the caTools package
```{r}
require(caTools)
data(mtcars)
dim(mtcars)
set.seed(2018) 
sample = sample.split(mtcars$mpg, SplitRatio = .75)
train = subset(mtcars, sample == TRUE)
test  = subset(mtcars, sample == FALSE)
dim(train)
dim(test)
```
### Approach# 4: using the createDataPartition() from the caret package
```{r}
library(caret)
dim(mtcars)
set.seed(2018)
splitIndex<-createDataPartition(y=mtcars$mpg,p=0.7,list=FALSE)
train<-mtcars[splitIndex,]
test<-mtcars[-splitIndex,]
dim(train)
dim(test)
```
